Rationality, Evolution \& the Environment

### 2015-10-04 paper/paper.tex
minor edits

A strategic game is a tuple $ G=\tuple{ N, (A_i , \pi_i)_{i \in N} }$ with $N$ a set of
players, $A_i$ a set of acts and $\pi_i$ a payoff function for each player $i \in N$. The
classical notion of rationality from Definition~\ref{def:rationality} also applies to
games. Let $A_{-i} = \bigtimes A_j$ for all $j \neq i$ be the set of all profiles of acts for
all players excepts $i$. Then player $i$ has a belief $\Delta(A_{-i})$ about the behavior of
his opponents.

@@

\todo[inline]{this section starts off presupposing to much formal stuff for a philosophy
  audience; we should first say why we look at games, then give the definition of a game, then
  define regret, then example, ...}

@@

\begin{example}[Coordination Game]
  Let us consider the coordination game in Figure \ref{coordgame1}. It is easy to see that
  $\text{reg}(I)=2$: when the other player chooses action $I$ the regret of $I$ is $0$ since
  $I$ is a best reply to $I$, and when the other player chooses $II$ the regret of $I$ is $2$
  since $\pi_i(II,II)-\pi_i(I,II)=2$. With the same reasoning, $\text{reg}(II)=1$. Two regret
  minimizing players would then choose action $II$, each getting a payoff of $2$. In this case
  regret minimization selects the Pareto-dominant Nash equilibrium. $ \medsquare $

@@

  \begin{subfigure}[b]{0.4\textwidth}
    \centering
    \begin{tabular}{ccc}
      \toprule
      & I & II \\
      \midrule
      I & 1;1 & 0;0 \\
      \hline 
      II & 0;0 & 2;2\\
      \bottomrule
    \end{tabular}
    \caption{Coordination game $G^C$}
    \label{coordgame1}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.4\textwidth}
    \centering
    \begin{tabular}{ccc}
      \toprule
      & I & II \\
      \midrule
      I & 0;0 & -2;-1 \\
      \hline 
      II & -1;-2 & 0;0\\
      \bottomrule
    \end{tabular}
    \caption{Regret representation $\tau^{reg}(G^C)$}
    \label{coordgame1reg}
  \end{subfigure}
  \caption{A coordination game (left) and its regret representation (right).}
    \label{coordgame1mainFig}
\end{figure}

>>>>

\subsection{Games} \label{sec:games}

A strategic game is a tuple $ G=\tuple{ N, (A_i , \pi_i)_{i \in N} }$ with $N$ a set of
players, $A_i$ a set of acts and $\pi_i$ a payoff function for each player $i \in N$. The
classical notion of rationality from Definition~\ref{def:rationality} also applies to
games. Let $A_{-i} = \bigtimes A_j$ for all $j \neq i$ be the set of all profiles of acts for
all players excepts $i$. If player $i$ has a belief $\mu_i \in \Delta(A_{-i})$ about the
behavior of his opponents, he chooses rationally only his choice $a^*_i$ maximizes expected
utility $\EU(a) = \sum_{a_{-i} \in A_{-i}} \mu_i \myts \pi_i(a_i, a_{-i})$.

We focus here on symmetric games with two players. A two-player game is symmetric if An example
is given in Figure~\ref{coordgame1}. \textcolor{red}{EXPAND}

\begin{itemize}
\item include here: minimax and choice given a belief
\end{itemize}

\todo[inline]{should we define NE, ESS/NSS here? dominance? we should include \emph{best
    reply}, because it is used in the definition of regret.}

@@

@@

\begin{example}[Coordination Game]
  Let us consider the coordination game in Figure \ref{coordgame1}. Figure~\ref{coordgame1reg}
  shows the pairwise regrets and the regrets of each row-player action. It is easy to see that
  $\text{reg}(I)=2$: when the other player chooses action $I$ the pairwise regret $reg(I,I)$ is
  $0$ since $I$ is a best reply to $I$, and when the other player chooses $II$ the pairwise
  regret $reg(I, II)$ is $2$ since $\pi_i(II,II)-\pi_i(I,II)=2$. With the same reasoning,
  $\text{reg}(II)=1$. Two regret minimizing players would then choose action $II$, each getting
  a payoff of $2$. In this case regret minimization selects the Pareto-dominant Nash
  equilibrium.\todo{explain Pareto-dominant NE; otherwise skip it?} $ \medsquare $

@@

  \begin{subfigure}[b]{0.3\textwidth}
    \centering
    \begin{tabular}{ccc}
      \toprule
      & I & II \\
      \midrule
      I & 1 & 0 \\
      II & 0 & 2\\
      \bottomrule
    \end{tabular}
    \caption{Coordination game $G^C$}
    \label{coordgame1}
  \end{subfigure}
  \hspace{1cm}
  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \begin{tabular}{cccc}
      \toprule
      & $reg( \cdot, I)$ & $reg(\cdot, II)$ & $reg(\cdot)$ \\
      \midrule
      I  & 0 & 2 & 2 \\ 
      II & 1 & 0 & 1\\
      \bottomrule
    \end{tabular}
    \caption{Pairwise regrets representation $\tau^{reg}(G^C)$}
    \label{coordgame1reg}
  \end{subfigure}
  \caption{A coordination game (left) and regrets associated with each row-player actions (right).}
    \label{coordgame1mainFig}
\end{figure}
